---
title: "Pittsburgh Steelers Database Marketing Analyst Assesment"
author: "Charlie Pallett"
output:
html_document:
df_print: paged
---

**Set Up**

Set up libraries and tools 

```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(ggimage)
library(RCurl)
library(tidyverse)
library(ggrepel)
library(plyr)
library(cluster)
library(factoextra)
library(rmarkdown)
library("writexl")
library(clustMixType)
```


Load Data

```{r}
data <-read_excel("steelers.xlsx")
data[is.na(data)] <- 0
```

Summary of Data

```{r}
summary(data)
```

**General Data Processing**

```{r}
#structure of data
str(data)

#total amuont of observations - 3000
length(data$id) 

#number of unqie user IDs in the data - 300
length(unique(data$id))
```
**Question #1**

1. Season ticket holder ticket utilization. Explain how you defined utilization.

Define Utilization: What is the each tickets end point?

Under the scope of this analysis, there are six possible ticket options for each game:

Flow chart: 
Active, Yes
Active, No
Forward, Yes
Forward, No
Resold, Yes
Resold, No

General breakdown?
Per person: How did they behave? 
Per game how were tickets utilized as a whole?
Idea: Pie Chart

Finding all outcomes different outcomes for tickets - totals & percentages 

```{r}
act_yes <- filter(data, ticket_action =="Active" , attended=="Y")
act_no <- filter(data, ticket_action =="Active" , attended=="N")
forward_yes <- filter(data, ticket_action =="Forward" , attended=="Y")
forward_no <- filter(data, ticket_action =="Forward" , attended=="N")
resold_yes <- filter(data, ticket_action =="Resold" , attended=="Y")
resold_no <- filter(data, ticket_action =="Resold" , attended=="N")

# How many of each occurrence? 
tally(act_yes) #512
tally(act_no) #516
tally(forward_yes) #493
tally(forward_no) #488
tally(resold_yes) #509
tally(resold_no) #482

# Percentage Breakdown
per1 = tally(act_yes)/length(data$id) # 0.1760667
per2 = tally(act_no)/length(data$id) # 0.172	
per3 = tally(forward_yes)/length(data$id) #0.1643333	
per4 = tally(forward_no)/length(data$id) #0.1626667	
per5 = tally(resold_yes)/length(data$id) #0.1696667
per6 = tally(resold_no)/length(data$id) #0.1606667	
```

**Key Facts**

What was most attended game?

A: Packers

Most resold game?

A: Browns 

Most forward but no resold game?

A: Jaguars 

```{r}
#Game by game breakdown vs. each team 
#Active, Yes
by_game1 <- act_yes %>% group_by(opponent)
df1 <- by_game1 %>% tally(sort = TRUE)
               
# Active, No
by_game2 <- act_no %>% group_by(opponent)
df2 <- by_game2 %>% tally(sort = TRUE)

# Forward, Yes
by_game3 <- forward_yes %>% group_by(opponent)
df3 <- by_game3 %>% tally(sort = TRUE)

# Forward, No
by_game4 <- forward_no %>% group_by(opponent)
df4 <- by_game4 %>% tally(sort = TRUE)

# Resold, Yes
by_game5 <- resold_yes %>% group_by(opponent)
df5 <- by_game5 %>% tally(sort = TRUE)

# Resold, No
by_game6 <- resold_no %>% group_by(opponent)
df6 <- by_game6 %>% tally(sort = TRUE)

#Group all outcomes by opponent 
df_list <- list(df1, df2, df3,df4,df5,df6)
teams <- df_list %>% reduce(full_join, by = 'opponent')
teams 
colnames(teams) <- c("opponent","act_yes", "act_no", "fwd_yes", "fwd_no", "res_yes","res_no")

#Data for team by team plot
teams
write_xlsx(teams,"~/gamebygame.xlsx")
```

Do people ultimately end up in the stadium? 

```{r}
#Active - 
a=tally(act_yes) #512
b=tally(forward_yes) #493
c=tally(resold_yes) #509
yes = a+b+c # 1514

#Not Active
d=tally(act_no) #516
e=tally(forward_no) #488
f=tally(resold_no) #482
no= d+e+f #1416

#Total Percentages Active vs Not Active
check=yes+no
yes/check #0.5046667
no/check #0.4953333	
```
Ticket Tiers 

```{r}
str(data)
#Ravens,Patriots at $150 
#Bengals, Seahawks, Vikings at $120
#Packers,Texans, Jags,Browns at $100
#Eagles at $80
```

**2. Season value per season ticket holder. Explain your value formula.**

Selected: Face values of originally purchased games + value of games sold on secondary markets / total amount of unique IDs

Alternative: Face values of originally purchased games + value of games sold on secondary markets + number of games ticket was forwarded / total amount of unique IDs

Note: 300 unique ID's 

```{r}
#Gross Values
x=sum(data$face_value) #342000
y=sum(data$secondary_price) #161457
z=tally(forward_yes) #493
gross1=x+y 
gross2=x+y+z

#Per Ticket Holder
val2=gross2/300 # 1679.833
val1=gross1/300 # 1678.19 --> going with this 
val1
```

Tier Calculations

```{r}
512/300 # Overall 
tt1=100/300 # Pats + Ravens (1)
tt2=151/300 #(2) 
tt3=204/300 #(3)
tt4=57/300#(4)
tt1
tt2
tt3
tt4
totchk = tt1+tt2+tt3+tt4
```
**3. Identify season ticket holder segments that should be contacted by a ticket rep. Explain why you are recommending a direct communication with these season ticket holders.**

Behavioural Cluster + Mediods

```{r}
datacopy <- data
datacopy
datacopy$opponent <- as.factor(datacopy$opponent)
datacopy$ticket_action <- as.factor(datacopy$ticket_action)
datacopy$attended <- as.factor(datacopy$attended)
unique(datacopy$opponent)
str(datacopy$ticket_action)
opponent = c('Patriots', 'Ravens','Vikings','Seahawks','Bengals','Packers','Texans','Jaguars','Browns','Eagles')
tier = c(1,1,2,2,2,3,3,3,3,4)
team_tier = cbind(opponent,tier)
team_tier
teamlist = list(team_tier)
clustdata <- teamlist%>%  reduce(full_join, by = opponent)
clustdata <- merge(datacopy,clustdata)

model_data <- select(clustdata,tier,face_value,ticket_action,attended,secondary_price)
model_data
clarafb <- clara(model_data,k=6, samples = 3000, pamLike = TRUE)
```

Fans who went to at least one game --> "Loyal"

```{r}
loyal <- act_yes %>% group_by(id)
loyal_count <- loyal %>% tally(sort = TRUE)
loyal_count
sum(loyal_count$n)

sum(loyal_count$n)
mean(loyal_count$n) #2.03

#majority of season ticket holder only go and scan to 1 game
hist(loyal_count$n)
```
Cancel candidates 

```{r}
data
cancel <- filter(data, ticket_action =="Active"  | ticket_action == "Forward", attended =="N")
can <- cancel %>% group_by(id)
can_count <- can %>% tally(sort = TRUE)
can_count
write_xlsx(can_count,"~/can_count.xlsx")
hist(can_count$n)

125/300 #4 and over --> 41%
```


Resellers 

```{r}
resellers2 <- filter(data, ticket_action =="Resold", attended =="N" | attended == "Y")
reso <- resellers2 %>% group_by(id)
res_count <- reso%>% tally(sort = TRUE)
res_count
write_xlsx(res_count,"~/res_count2.xlsx")
65/300 # over five ---> 22%  
```

Calculations

```{r}
g1=count(loyal_count$n == 1) #101
g2=count(loyal_count$n == 2) #81
g3=count(loyal_count$n == 3) #51
g4=count(loyal_count$n == 4) #13
g5=count(loyal_count$n == 5) #5
g6=count(loyal_count$n == 6) #0
g7=count(loyal_count$n == 7) #0
g8=count(loyal_count$n == 8) #0
g9=count(loyal_count$n == 9) #1
g10=count(loyal_count$n == 10) #1
101+81
51+13+5+1+1
81+51+13+5+2
152+101
```

Save Resale to use in plot

```{r}
#resell
res_guys <- resold_yes %>% group_by(id)
res_count <- res_guys  %>% tally(sort = TRUE)
res_count
hist(res_count$n)
write_xlsx(res_count,"~/rescount.xlsx")
```

Save Loyal Fans to Use in Plot

```{r}
sort <- data %>% group_by(id)
dfsort <- sort %>% tally(sort = TRUE)
dfsort

loyal <- act_yes %>% group_by(id)
loyal_count <- loyal %>% tally(sort = TRUE)
loyal_count
write_xlsx(loyal_count,"~/loyalcount.xlsx")
```

More Calculations 

```{r}
#percentage that went to at least one game
253/300 #84.3% 

#percentatage of fans that go to more than 6 game s
(2/300)*100 #0.67% 

#percentage of fans that attend 4 or less games #82% 
lessfour = 101+81+51+13
x=lessfour/300
x
```

Summary of Cluster Totals 

```{r}
summary(as.factor(clarafb$clustering))
```

Mediods from Clusters

```{r}
clarafb$medoids
```

Check Results
```{r}
summary <- cbind(model_data, cluster = clarafb$cluster)
results_check <- head(summary, n = 50)
paged_table(results_check)
results_check
```

**Rough Clusters Chart** 

```{r}
clustplotfb = fviz_cluster(clarafb, choose.vars = NULL, stand = TRUE,
  axes = c(1, 2), geom = c("point"), repel = FALSE,
  show.clust.cent = TRUE, ellipse = TRUE, ellipse.type = "convex",
  ellipse.level = 0.95, ellipse.alpha = 0.2, shape = NULL,
  pointsize = 1.5, labelsize = 12, main = "Cluster plot",
  xlab = NULL, ylab = NULL, outlier.color = "black",
  outlier.shape = 19, outlier.pointsize = pointsize,
  outlier.labelsize = labelsize, ggtheme = theme_grey())
clustplotfb
```
```{r}
#colMeans(new.segment.data[clara.res2$clustering == 1,colnames(clara.res2$medoids)], na.rm = TRUE)
```

What are the proportions for each ticket price? 

$150 > 600
$120 > 900
$100 > 1200
$80 > 300 


**4. Secondary lift (left?) over face value for each game**

Idea: Game by game show the average increase in secondary market vs. first game

```{r}
data2 <-read_excel("steelers2.xlsx")
#data2
```

What percentage of tickets were re-sold? 

```{r}
resaleper = 991/3000 #0.3303334 --> 33%
resaleper
#0.1606667+0.1696667 resold values above
```

Breakdown of re-sell actions

```{r}
482/991 # 0.4863774 --> resold and didn't scan (NO)
509/991 #0.5136226 --> resold and didn't scan (YES)
```

Most frequently re-sold game

```{r}
res1 <- data2 %>% group_by(opponent)
res1 <- res1 %>% tally(sort = TRUE)
res1
```

Total Re-sale per opponent 

```{r}
resale_tot = data2 %>%                            
  group_by(opponent) %>%                         
  summarise_at(vars(secondary_price),              
               list(secondary_price = sum))
resale_tot
```

Average re-sale price per opponent 

```{r}
#Bengals
14738/109	#135.211

#Browns 
15182/106	#143.2264

#Eagles 
7508/105 #71.50476

#Jaguars
13907/102 #136.3431


#Packers
14750/102 #136.3431

#Patriots 
23205/101	#229.7525

#Ravens 
25370/97 #261.5464

#Seahwawks
14731/90 #163.6778

#Texans 
15863/90 #176.2556

#Vikings
16203/89 #182.0562
```

Compare this to average face value price

Produce Bar Graph showing game by game 


